---
title: "cumulativedata"
output:
  html_document: default
  pdf_document: default
date: "2024-02-15"
---
CUMULATIVE 
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
cumulativeraw = read.csv2("/Users/oliviacandelaria/Documents/T4S/T4S Survey Data - Cumulative.csv",sep=',')
View(cumulativeraw)
cumulative = cumulativeraw
```

Logical Variable for Initial Learn and Learn Numeric 
```{r}
cumulative$Initial.Learn <- as.logical(ifelse(is.na(cumulative$Initial.Learn), 
                                          NA, cumulative$Initial.Learn == "Yes"))
cumulative['learn_numeric_initial'] <-as.numeric(cumulative$Initial.Learn) + 
  is.na(cumulative$Initial.Learn)

View(cumulative)
```
# cumulative now has logical variable for Initial.Learn and
# new variable learn_numeric that converted logical to 0/1


Creates DF with Averages Column
```{r}
grouped_sums <- cumulative %>%
  group_by(Gender) %>%
  summarize(sum = sum(na.omit(learn_numeric_initial)))
grouped_sums
cumulative['LearnAverage']=NA

for (i in 1:nrow(cumulative)) {
  if (!is.na(cumulative$Initial.Learn[i]) && !is.na(cumulative$Gender[i])) {
    if (cumulative$Initial.Learn[i] & cumulative$Gender[i] == "M") {
      cumulative$LearnAverage[i] <- grouped_sums$sum[2] / sum(cumulative$Gender=="M")
    } else if (!cumulative$Initial.Learn[i] & cumulative$Gender[i] == "M") {
      cumulative$LearnAverage[i] <- (sum(cumulative$Gender=="M") - grouped_sums$sum[2]) / sum(cumulative$Gender=="M")
    } else if (cumulative$Initial.Learn[i] & cumulative$Gender[i] == "F") {
      cumulative$LearnAverage[i] <- grouped_sums$sum[1] / sum(cumulative$Gender=="F")
    } else {
      cumulative$LearnAverage[i] <- (sum(cumulative$Gender=="F") - grouped_sums$sum[1]) / sum(cumulative$Gender=="F")
    }
  }
}
View(cumulative)
```

Gender Difference in Willingness to Learn
```{r}
filtered_data <- cumulative[!is.na(cumulative$Initial.Learn), ]

ggplot(filtered_data,aes(x=Initial.Learn,y=LearnAverage,fill = Gender)) +
  geom_bar(stat= 'identity',position = "dodge")+ labs(title = "Gender Difference in Willingness to Learn")+
  scale_x_discrete(labels = c("No","Yes")) +
  labs(x="Do you like to learn?")+
  #annotate("text", x = 1.75, y = .65, label = "0.673")+ annotate("text", x = 2.25, y = .6, label = "0.625")+
  #annotate("text", x = 0.75, y = .3, label = "0.327")+ annotate("text", x = 1.25, y = .35, label = "0.375")+
  theme_minimal()
```

Average Confidence Levels by Gender
```{r}
#View(cumulative%>%group_by(Gender,average))

cumulative2 <- (cumulative) %>% 
pivot_longer(cols = c("Initial.Public", "Initial.Small", "Initial.Adults"), 
               names_to = "ConfLevels", 
               values_to = "Level")
#View((cumulative2))

cumulative3 <- ((cumulative2))%>% 
  group_by(ConfLevels, Gender) %>% 
  mutate(LevelAverage = mean(Level, na.rm=T))

#View(cumulative3[,c(1,17)])

ggplot(cumulative3,aes(x = ConfLevels, y = LevelAverage , fill = Gender))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Gender", x = "Variables",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Gender"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()

```

Creates Age Range Variable
```{r}
cumulative['agerange'] <- NA

for(i in 1:nrow(cumulative)){
  if (!is.na(cumulative$Age[i])) {
    if(cumulative$Age[i]<9){
      cumulative$agerange[i] = "6-8"
    }else if(9<=cumulative$Age[i] & cumulative$Age[i]<12){
      cumulative$agerange[i] = "9-11"
    }else{
      cumulative$agerange[i] = "12-14"
    }
  }
}
```

Average Public Speaking Confidence by Age
```{r}
data_age <- cumulative %>%
  group_by(agerange)%>%
  mutate(PublicAverage = mean(Initial.Public))

level_order <- c("6-8", "9-11", "12-14")

ggplot(na.omit(data_age)) + geom_bar(aes(x = factor(agerange, level = level_order),
       y = PublicAverage, fill = factor(agerange, level = level_order)), stat = "identity", position = "dodge", 
       width = 0.5) + theme_minimal() + guides(fill=guide_legend(title="Age"))+
       labs(title = "Average Public Speaking Confidence by Age", x="Age",
            y="Public Speaking Confidence")
```
Average Small Group Confidence by Age
```{r}
data_age <- data_age %>%
  group_by(agerange)%>%
  mutate(SmallAverage = mean(na.omit(Initial.Small)))

ggplot(na.omit(data_age)) + geom_bar(aes(x = factor(agerange, level = level_order),
       y = SmallAverage, fill = factor(agerange, level = level_order)), stat = "identity", position = "dodge", 
       width = 0.5) + theme_minimal() + guides(fill=guide_legend(title="Age"))+
       labs(title = "Average Small Group Confidence by Age", x="Age",
            y="Small Group Confidence")
```
Average Confidence Speaking to Adults by Age
```{r}
data_age <- data_age %>%
  group_by(agerange)%>%
  mutate(AdultAverage = mean(na.omit(Initial.Adults)))

ggplot(na.omit(data_age)) + geom_bar(aes(x = factor(agerange, level = level_order),
       y = AdultAverage, fill = factor(agerange, level = level_order)), stat = "identity", position = "dodge", 
       width = 0.5) + theme_minimal() + guides(fill=guide_legend(title="Age"))+
       labs(title = "Average Confidence Speaking to Adults by Age", x="Age",
            y="Confidence with Adults")
```
Average Confidence Levels by Age
```{r}
cumulative3['agerange'] <- NA

for(i in 1:nrow(cumulative3)){
  if (!is.na(cumulative3$Age[i])) {
    if(cumulative3$Age[i]<9){
      cumulative3$agerange[i] = "6-8"
    }else if(9<=cumulative3$Age[i] & cumulative3$Age[i]<12){
      cumulative3$agerange[i] = "9-11"
    }else{
      cumulative3$agerange[i] = "12-14"
    }
  }
}

cumulative4 <- cumulative3 %>% 
  group_by(ConfLevels, agerange) %>% 
  mutate(AgeAverage = mean(Level))

cumulative4_filtered <- cumulative4 %>% filter(!is.na(agerange))


ggplot(cumulative4_filtered,aes(x = ConfLevels, y = AgeAverage , fill =factor(agerange, level = level_order)))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Age", x = "Variables",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Age"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()
```

```{r}
summary(cumulative)
head(cumulative)

sum(na.omit(cumulative$agerange == "6-8"))
sum(na.omit(cumulative$agerange == "9-11"))
sum(na.omit(cumulative$agerange == "12-14"))
sum(cumulative$Gender == "M")
sum(cumulative$Gender == "F")

```



H0: mean confidence with adults 6-8 = mean confidence with adults 12-14
H1: mean confidence with adults 6-8 != mean confidence with adults 12-14
a = 0.05
```{r}
cumulative_subset <- cumulative[cumulative$agerange %in% c("6-8", "12-14"), ]

# Extract Initial.Adults for each age range
initial_adults_6_8 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "6-8"]
initial_adults_12_14 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "12-14"]

# Perform two-sample t-test
t_test_result <- t.test(initial_adults_6_8, initial_adults_12_14)

# Print the result
print(t_test_result)

# fail to reject the null, our initial hypothesis is right and there is no sig difference
```
p-value = 0.0516 so it is low enough to reject the null hypothesis that 
mean Initial.Adults is equal for 6-8 and 12-14
This means that there is significant difference between the grade groups

```{r}
cumulative_subset <- cumulative[cumulative$agerange %in% c("6-8", "9-11"), ]

# Extract Initial.Adults for each age range
initial_adults_6_8 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "6-8"]
initial_adults_9_11 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "9-11"]

# Perform two-sample t-test
t_test_result <- t.test(initial_adults_6_8, initial_adults_9_11)

# Print the result
print(t_test_result)

```

```{r}
cumulative_subset <- cumulative[cumulative$agerange %in% c("9-11", "12-14"), ]

# Extract Initial.Adults for each age range
initial_adults_9_11 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "9-11"]
initial_adults_12_14 <- cumulative_subset$Initial.Adults[cumulative_subset$agerange == "12-14"]

# Perform two-sample t-test
t_test_result <- t.test(initial_adults_9_11, initial_adults_12_14)

# Print the result
print(t_test_result)
```

Count Number of Black Students
```{r}
sum(grepl('Black',cumulative$Race))
```

Cleaning Race Variable
```{r}
#Changes all Hispanic/black students to just black or just white
RaceData <- cumulative2
for(i in 1:nrow(cumulative2)){
  if(grepl('/Black', cumulative2$Race[i])){
    RaceData$Race[i] <- 'Black'
  } else if(grepl('/White', cumulative2$Race[i])){
    RaceData$Race[i] <- 'White'
  }
}

sum(grepl('Hispanic', cumulative$Race))
sum(grepl('Black', cumulative$Race))
sum(grepl('White', cumulative$Race))
sum(grepl('Hispanic/White', cumulative$Race))
sum(grepl('Hispanic/Black', cumulative$Race))


```

Average Confidence Levels by Ethnicity
```{r}
cumulative_race <- cumulative2 %>% 
  group_by(ConfLevels, Race) %>% 
  mutate(RaceAverage = mean(Level, na.rm = T))
#levels(factor(cumulative$Race))

race_vector <- c('Asian', 'Black', 'Hispanic', 'White')
cumulative_race_filtered <- cumulative_race[cumulative_race$Race %in% race_vector, ]
cumulative_race_filtered <- cumulative_race_filtered %>% filter(!is.na(Race))

#View(cumulative_race_filtered)
ggplot(cumulative_race_filtered, aes(x = ConfLevels, y = RaceAverage , fill =Race))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Ethnicity", x = "",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Ethnicity"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()
```

Average Confidence Levels by Ethnicity (without the 1 asian student)
- also put all Hispanic/Black into black category and Hispanic/White into white
```{r}
cumulative_race <- RaceData %>% 
  group_by(ConfLevels, Race) %>% 
  mutate(RaceAverage = mean(Level, na.rm = T))
#levels(factor(cumulative$Race))

race_vector <- c('Black', 'Hispanic', 'White')
cumulative_race_filtered <- cumulative_race[cumulative_race$Race %in% race_vector, ]
cumulative_race_filtered <- cumulative_race_filtered %>% filter(!is.na(Race))

#View(cumulative_race_filtered)
ggplot(cumulative_race_filtered, aes(x = ConfLevels, y = RaceAverage , fill =Race))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Ethnicity", x = "",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Ethnicity"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()
```

Biracial Graph
```{r}
RaceData$Race <- ifelse(cumulative2$Race %in% c("Hispanic/Black", "Hispanic/White"), "Biracial", cumulative2$Race)
View(RaceData)
cumulative_race <- RaceData %>% 
  group_by(ConfLevels, Race) %>% 
  mutate(RaceAverage = mean(Level, na.rm = T))
#levels(factor(cumulative$Race))


race_vector <- c('Black', 'Hispanic', 'White', 'Biracial')
cumulative_race_filtered <- cumulative_race[cumulative_race$Race %in% race_vector, ]
cumulative_race_filtered <- cumulative_race_filtered %>% filter(!is.na(Race))

#View(cumulative_race_filtered)
ggplot(cumulative_race_filtered, aes(x = ConfLevels, y = RaceAverage , fill =Race))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Ethnicity", x = "",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Ethnicity"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()
```


Race graph 
```{r}
race_split_data = cumulative
for(i in 1:nrow(race_split_data)){
  if(grepl('/',race_split_data$Race[i])){
    race_split <- strsplit(as.character(race_split_data$Race[i]), "/")[[1]]
    if(length(race_split)<3){
      race1 <- race_split[1]
      race2 <- race_split[2]
      race_split_data$Race[i] = race1
      row_to_duplicate <- race_split_data[i, ]
      row_to_duplicate$Race <- race2
      race_split_data <- rbind(race_split_data, row_to_duplicate)
    }else{
      race1 <- race_split[1]
      race2 <- race_split[2]
      race3 <- race_split[3]
      race_split_data$Race[i] = race1
      
      row_to_duplicate1 <- race_split_data[i, ]
      row_to_duplicate1$Race <- race2
      race_split_data <- rbind(race_split_data, row_to_duplicate1)
      
      row_to_duplicate2 <- race_split_data[i, ]
      row_to_duplicate2$Race <- race3
      race_split_data <- rbind(race_split_data, row_to_duplicate2)
    }
  }
}
View(race_split_longer)

race_split_longer <- (race_split_data) %>% 
pivot_longer(cols = c("Initial.Public", "Initial.Small", "Initial.Adults"), 
               names_to = "ConfLevels", 
               values_to = "Level")

race_split_longer <- ((race_split_longer))%>% 
  group_by(ConfLevels, Race) %>% 
  mutate(LevelAverage = mean(Level, na.rm=T))

race_split_longer <- race_split_longer %>% filter(!is.na(Race))
race_split_longer <- race_split_longer[race_split_longer$Race %in% race_vector, ]


ggplot(race_split_longer, aes(x = ConfLevels, y = LevelAverage , fill =Race))+
  geom_bar(stat = "identity", position = "dodge",  width = .5)+
  labs(title = "Average Confidence Levels by Ethnicity", x = "",
       y = "Average Level of Confidence")+
  guides(fill=guide_legend(title="Ethnicity"))+
  scale_x_discrete(labels=c("Initial.Public"="Public Speaking", "Initial.Adults" =
        "Speaking with Adults", "Initial.Small"="Speaking in a Small Group"))+
  theme_minimal()
```





