---
title: "Non Monroe Data"
output: html_document
date: "2024-03-21"
---
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(gridExtra)
```

Load Data Sets
```{r}
cumulativeraw = read.csv2("/Users/oliviacandelaria/Documents/T4S/DATA/T4S(Cumulative).csv",sep=',')
View(cumulativeraw)
```
```{r}
longer = read.csv2("/Users/oliviacandelaria/Documents/T4S/DATA/T4S(Longer).csv",sep=',')
View(longer)
```

Create NON Monroe Data
```{r}
data <- longer%>%
  filter(!grepl('Monroe',School))
#View(data)
```

Split multiracial students into new data set
```{r}
race_split_data = data
for(i in 1:nrow(data)){
  if(grepl('/',race_split_data$Race[i])){
    race_split <- strsplit(as.character(race_split_data$Race[i]), "/")[[1]]
    if(length(race_split)<3){
      race1 <- race_split[1]
      race2 <- race_split[2]
      race_split_data$Race[i] = race1
      row_to_duplicate <- race_split_data[i, ]
      row_to_duplicate$Race <- race2
      race_split_data <- rbind(race_split_data, row_to_duplicate)
    }else{
      race1 <- race_split[1]
      race2 <- race_split[2]
      race3 <- race_split[3]
      race_split_data$Race[i] = race1
      
      row_to_duplicate1 <- race_split_data[i, ]
      row_to_duplicate1$Race <- race2
      race_split_data <- rbind(race_split_data, row_to_duplicate1)
      
      row_to_duplicate2 <- race_split_data[i, ]
      row_to_duplicate2$Race <- race3
      race_split_data <- rbind(race_split_data, row_to_duplicate2)
    }
  }
}
#View(race_split_data)

race_vector <- c('Black', 'Hispanic', 'White')
cumulative_race_filtered <- race_split_data[race_split_data$Race %in% race_vector, ]
```

Create Age Range column in data set
```{r}
data['agerange'] <- NA

for(i in 1:nrow(data)){
  if (!is.na(data$Age[i])) {
    if(data$Age[i]<9){
      data$agerange[i] = "6-8"
    }else if(9<=data$Age[i] & data$Age[i]<12){
      data$agerange[i] = "9-11"
    }else{
      data$agerange[i] = "12-14"
    }
  }
}
#View(data)
```

Plots:

Public Speaking by Race
```{r}
grouped_data <- cumulative_race_filtered %>%
  group_by(Status, Race)
summary_data <- (grouped_data) %>%
  summarise(mean_value = mean(Public, na.rm=T))
summary_data

desired_order <- c("Initial","Final")

public_plot <- ggplot(summary_data, aes(x=Race,y=mean_value,fill = factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(title='',subtitle = 'Public Speaking', 
       x = "", y = "Average Level of Confidence")+
  guides(fill=guide_legend(title=""))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,5))
```

Small Group by Race
```{r}
grouped_data <- cumulative_race_filtered %>%
  group_by(Status, Race)
summary_data <- (grouped_data) %>%
  summarise(mean_value = mean(Small, na.rm=T))
summary_data

desired_order <- c("Initial","Final")

small_plot <- ggplot(summary_data, aes(x=Race,y=mean_value,fill = factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(title = "",subtitle = 'Small Group', 
       x = "", y = "")+
  guides(fill=guide_legend(title=""))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,5))
```

Adult by Race
```{r}
grouped_data <- cumulative_race_filtered %>%
  group_by(Status, Race)
summary_data <- (grouped_data) %>%
  summarise(mean_value = mean(Adult, na.rm=T))
summary_data

desired_order <- c("Initial","Final")

adult_plot <- ggplot(summary_data, aes(x=Race,y=mean_value,fill = factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(title = "",subtitle = 'Speaking with Adults', 
       x = "", y = "")+ guides(fill=guide_legend(title=""))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,5))
```

Arrange plots side by side
```{r}
combined_plot <- grid.arrange(public_plot, small_plot, adult_plot, ncol = 3, 
                              top = 'Confidence Levels by Ethnicity\nInitial vs. Final')
combined_plot
```

Do you like to learn by race: 
```{r}
proportions <- cumulative_race_filtered %>%
  group_by(Status, Race, Learn) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count))%>%
  filter( Learn != '')

desired_order<-c('Initial','Final')
learn_plot<-ggplot(proportions, aes(x=Learn,y=proportion,fill = factor(Status, desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  scale_y_continuous(limits=c(0,1))+ facet_wrap(~Race)+
  labs(title='Do You Like to Learn by Ethnicity\nInitial vs. Final',x='')+
  theme(legend.position = 'top')

learn_plot
```


Age Range Public Speaking
```{r}
grouped_data <- data %>%
  group_by(Status, agerange)%>%
  summarise(mean_value = mean(Public, na.rm=T))%>%
  filter(agerange != '')
grouped_data

desired_order <- c("Initial","Final")
age_order <-c('6-8','9-11','12-14')


public_plot <- ggplot(grouped_data, aes(x=factor(agerange,level=age_order),y=mean_value,fill = 
                                          factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(subtitle = 'Public Speaking', 
       x = "", y = "")+
  guides(fill=guide_legend(title=""))+
  scale_y_continuous(limits=c(0,5))+  theme(legend.position = "none")
public_plot
```

Age Range Small Group
```{r}
grouped_data <- data %>%
  group_by(Status, agerange)%>%
  summarise(mean_value = mean(Small, na.rm=T))%>%
  filter(agerange != '')
grouped_data

desired_order <- c("Initial","Final")
age_order <-c('6-8','9-11','12-14')

small_plot <- ggplot(grouped_data, aes(x=factor(agerange,level=age_order),y=mean_value,fill =
                                          factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(subtitle = 'Small Group', 
       x = "Age Range", y = "")+
  guides(fill=guide_legend(title=""))+
  scale_y_continuous(limits=c(0,5))+  theme(legend.position = "none")
small_plot
```

Age Range Adults
```{r}
grouped_data <- data %>%
  group_by(Status, agerange)%>%
  summarise(mean_value = mean(Adult, na.rm=T))%>%
  filter(agerange != '')
grouped_data

desired_order <- c("Initial","Final")
age_order <-c('6-8','9-11','12-14')

adult_plot <- ggplot(grouped_data, aes(x=factor(agerange,level=age_order),y=mean_value,fill =
                                          factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  labs(subtitle = 'Speaking to Adults', 
       x = "", y = "")+
  guides(fill=guide_legend(title=""))+
  scale_y_continuous(limits=c(0,5))+  theme(legend.position = "none") 
adult_plot
```

Combined age range
```{r}
combined_plot <- grid.arrange(public_plot, small_plot, adult_plot, ncol = 3, 
                              top = 'Confidence Levels by Age Range\nInitial vs. Final')
combined_plot
```

Combined Confidence Levels Age Range
```{r}
data_levels <- data %>% 
  pivot_longer(cols = c("Public", "Small", "Adult"), 
               names_to = "ConfLevels", 
               values_to = "Level")%>%
  group_by(Status, agerange, ConfLevels)%>%
  summarise(mean_value = mean(Level, na.rm=T))%>%
  filter(agerange!='')
data_levels

ggplot(data_levels, aes(x=factor(agerange,level=age_order),y=mean_value,fill =
                                          factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  scale_y_continuous(limits=c(0,5)) +facet_wrap(~factor(ConfLevels,c("Public", "Small", "Adult") ))+
  guides(fill=guide_legend(title=""))+labs(title='Initial/Final Confidence Levels',
                                           subtitle='Non-Monroe Schools',x='Age Range')

```

Do you like to learn by age group
```{r}
proportions <- data %>%
  group_by(Status, Gender, Learn) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count))%>%
  filter( Learn != '')


learn_plot<-ggplot(proportions, aes(x=Learn,y=proportion,fill = factor(Status,desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  scale_y_continuous(limits=c(0,1))+ facet_wrap(~Gender)+
  labs(title='Do You Like to Learn by Gender\nInitial vs. Final',x='')+
  guides(fill=guide_legend(title=''))

learn_plot
```

Confidence levels by gender
```{r}
data_levels <- data %>% 
  pivot_longer(cols = c("Public", "Small", "Adult"), 
               names_to = "ConfLevels", 
               values_to = "Level")%>%
  group_by(Status, Gender, ConfLevels)%>%
  summarise(mean_value = mean(Level, na.rm=T))%>%
  filter(Gender!='')
data_levels
desired_order <- c('Initial','Final')

ggplot(data_levels, aes(x=factor(Gender),y=mean_value,fill = factor(Status,level=desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  scale_y_continuous(limits=c(0,5)) +
  
  facet_wrap(~factor(ConfLevels,c("Public", "Small", 'Adult'),labels = c('Public Speaking', 'Small Group', 'Speaking to Adults') ))+
  
  guides(fill=guide_legend(title=""))+
  labs(title='Initial/Final Confidence Levels by Gender', subtitle='Week-long Program',
       x='Gender', y= 'Average Reported Confidence\n')
```


Check for significance in Gender Public Speaking:
t-test for public speaking F vs M
h0: mean of public F = mean of public M
h1: mean of public F < mean of public M
```{r}
#Get 2 vectors, one for girls one for boys
public_girl <- data %>% filter(Gender=="F") %>% select(Public)
public_boy <- data %>% filter(Gender=="M") %>% select(Public)

t.test(public_girl,public_boy)
#p-value = 0.00125, indicates statistical difference in means
#this means that mean of public girls is significantly less than boys
#(combined initial and final surveys)
```

Overall Like to Learn
```{r}
proportions <- cumulative_race_filtered %>%
  group_by(Status, Learn) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count))%>%
  filter( Learn != '')

desired_order<-c('Initial','Final')
learn_plot<-ggplot(proportions, aes(x=Learn,y=proportion,fill = factor(Status, desired_order)))+
  geom_bar(stat= 'identity', position='dodge', width = 0.5)+
  scale_y_continuous(limits=c(0,1))+ 
  labs(title='Do You Like to Learn \nInitial vs. Final',x='')+
  theme(legend.position = 'top')

learn_plot
```



